Calibrating the pixel-level Kepler imaging data with a causal data-driven model \\

### 2015-03-09 documents/cpm.tex
reference and bibtex

@@

In general, there are many reasons for apparent photometric variability in a \Kepler\ source.
There is intrinsic stellar variability,
  which is of interest to some and a nuisance to others.
There is also variability of overlapping fainter stars;
  that is, confusion noise combined with variability of the confusing sources.
There are small changes in spacecraft pointing,
  which leads to slightly different illumination of the focal-plane pixels,
  and thus different sensitivity to variations in the device
There are also \emph{intra-pixel} sensitivity variations that can contribute (\todo{cite white paper; Spitzer results}).
There are small changes in spacecraft temperature,
  which lead to point-spread function (PSF) and differential (across the focal plane) pointing changes.
These also lead to changes in pixel and intra-pixel illumination.
Stellar proper motion, geometric parallaxes, and differential stellar aberration as the spacecraft orbits all do more of the same.
There is electronic cross-talk between detectors and charge-transfer inefficiency;
  these can effectively transfer variability from one source to another.
There are additional electronics effects like ``rolling bands'' that put additional features into lightcurves.
There are also probably changes to the detector sensitivity with temperature and time,
  possibly illimunation-history effects,
  and possibly sources of variability not yet considered.
The remarkable thing about \Kepler\ is that it is trying to measure stars at a level of precision
  much higher than ever previously attempted;
  new effects really \emph{must} appear at some point.
In \figurename~\ref{fig:pixelpatch}, we show the pixel-level variability in the \Kepler\ data
  near one bright star that shows minimal intrinsic variability;
  this figure highlights the spacecraft-induced effects.

@@

One such method---one that is designed to describe or model spacecraft-induced problems
  but \emph{not} interfere with measurements of stellar variability---%
  is the \Kepler\ Presearch (\sic) Data Conditioning (PDC) method (\todo{citation}).
The PDC builds on the idea that systematic errors have a temporal structure that can be extracted from ancillary quantities. 
%This seems the original paper on PDC:
%[PDC1] https://kepler.nasa.gov/science/ForScientists/papersAndDocumentation/SOCpapers/7740-67-pdc-jtwicken-copyright.pdf
%
%This describes the later version of PDC:
%[PDC2] http://arxiv.org/pdf/1203.1382v1.pdf
%[PDC3] http://arxiv.org/pdf/1203.1383v1.pdf

In the first PDC \cite{PDC1}, removal of systematic errors was performed based on correlations with a set of ancillary engineering data. These data include the temperatures at the local detector electronics below the CCD array, and polynomials describing the centroid motion of the targets from PA.

In \cite{PDC2,PDC3}, filtered light curves of other stars are used instead. The light curve of a target star is projected on the leading eight principal components of these light curves. This projection is then removed from the star's light curve. The principal components are computed from a set of relatively quiet stars close in position and magnitude. This was done since it was observed that otherwise, too much of the stellar variability was being removed.

From the point of view of the present paper, one can only guarantee that {\em no} stellar variability is removed if the stars do not depend on time at all. If both the target star signal as well as some of the other stars do depend on time, then time acts as a confounder influencing both, and thus the true signals are not statistically independent, and thus the target signal may be compromised. We suspect that this effect is limited in \cite{PDC2,PDC3} since only eight principal components are used, which limits the capacity of the model.

@@

\begin{figure}[htb]
\centering
\begin{subfigure}[htb]{0.48\columnwidth}
\includegraphics[width=\columnwidth]{5088536-5}
\caption{}
\end{subfigure}%
\hfill
\begin{subfigure}[htb]{0.48\columnwidth}
\includegraphics[width=\columnwidth]{5949551-5}
\caption{}
\end{subfigure}
\caption{\label{ccd} Stars on the same CCD share systematic errors. 
The two panels show pixel fluxes (brightnesses) for two stars: (a) KIC 5088536, (b) KIC 5949551; 
here, KIC stands for Kepler Input Catalog. Both stars lie on the same CCD (see Fig.~\ref{fig1}), 
but far enough apart such that there is no stray light from one affecting the other. 
Each panel shows the pixels contributing to the respective star. 
Note that there exist similar trends in these two stars, caused by systematic errors. 
%The idea of our approach is to predict each pixel from a set of pixels from other stars in order to remove those systematics.
}
\end{figure}

@@

>>>>

\bibliographystyle{apj}

@@

In general, there are many reasons for apparent photometric variability in a \Kepler\ source.
There is intrinsic stellar variability,
  which is of interest to some and a nuisance to others.
There is also variability of overlapping fainter stars;
  that is, confusion noise combined with variability of the confusing sources.
There are small changes in spacecraft pointing,
  which leads to slightly different illumination of the focal-plane pixels,
  and thus different sensitivity to variations in the device
There are also \emph{intra-pixel} sensitivity variations that can contribute \citep{whitepaper} (\todo{cite white paper; Spitzer results}).
There are small changes in spacecraft temperature,
  which lead to point-spread function (PSF) and differential (across the focal plane) pointing changes.
These also lead to changes in pixel and intra-pixel illumination.
Stellar proper motion, geometric parallaxes, and differential stellar aberration as the spacecraft orbits all do more of the same.
There is electronic cross-talk between detectors and charge-transfer inefficiency;
  these can effectively transfer variability from one source to another.
There are additional electronics effects like ``rolling bands'' that put additional features into lightcurves.
There are also probably changes to the detector sensitivity with temperature and time,
  possibly illimunation-history effects,
  and possibly sources of variability not yet considered.
The remarkable thing about \Kepler\ is that it is trying to measure stars at a level of precision
  much higher than ever previously attempted;
  new effects really \emph{must} appear at some point.
In \figurename~\ref{ccd}, we show the pixel-level variability in the \Kepler\ data
  near one bright star that shows minimal intrinsic variability;
  this figure highlights the spacecraft-induced effects.

@@

One such method---one that is designed to describe or model spacecraft-induced problems
  but \emph{not} interfere with measurements of stellar variability---%
  is the \Kepler\ Presearch (\sic) Data Conditioning (PDC) method \citep{pdc1}, (\todo{citation}).
The PDC builds on the idea that systematic errors have a temporal structure that can be extracted from ancillary quantities. 
%This seems the original paper on PDC:
%[PDC1] https://kepler.nasa.gov/science/ForScientists/papersAndDocumentation/SOCpapers/7740-67-pdc-jtwicken-copyright.pdf
%
%This describes the later version of PDC:
%[PDC2] http://arxiv.org/pdf/1203.1382v1.pdf
%[PDC3] http://arxiv.org/pdf/1203.1383v1.pdf

In the first PDC \cite{pdc1}, removal of systematic errors was performed based on correlations with a set of ancillary engineering data. These data include the temperatures at the local detector electronics below the CCD array, and polynomials describing the centroid motion of the targets from PA.

In \cite{pdc2,pdc3}, filtered light curves of other stars are used instead. The light curve of a target star is projected on the leading eight principal components of these light curves. This projection is then removed from the star's light curve. The principal components are computed from a set of relatively quiet stars close in position and magnitude. This was done since it was observed that otherwise, too much of the stellar variability was being removed.

From the point of view of the present paper, one can only guarantee that {\em no} stellar variability is removed if the stars do not depend on time at all. If both the target star signal as well as some of the other stars do depend on time, then time acts as a confounder influencing both, and thus the true signals are not statistically independent, and thus the target signal may be compromised. We suspect that this effect is limited in \cite{pdc2,pdc3} since only eight principal components are used, which limits the capacity of the model.

@@

\begin{figure}[htb]
\centering
\begin{subfigure}[htb]{0.48\columnwidth}
\includegraphics[width=\columnwidth]{5088536-5}
\caption{}
\end{subfigure}%
\hfill
\begin{subfigure}[htb]{0.48\columnwidth}
\includegraphics[width=\columnwidth]{5949551-5}
\caption{}
\end{subfigure}
\caption{\label{ccd} Stars on the same CCD share systematic errors. 
The two panels show pixel fluxes (brightnesses) for two stars: (a) KIC 5088536, (b) KIC 5949551; 
here, KIC stands for Kepler Input Catalog. Both stars lie on the same CCD, 
but far enough apart such that there is no stray light from one affecting the other. 
Each panel shows the pixels contributing to the respective star. 
Note that there exist similar trends in these two stars, caused by systematic errors. 
%The idea of our approach is to predict each pixel from a set of pixels from other stars in order to remove those systematics.
}
\end{figure}

@@

\bibliography{cpm}
